---
title: ""
self_contained: yes
runtime: shiny
---

<br>

## Biological data

```{r setup, include = FALSE}
# globals
library(knitr)
opts_chunk$set(echo = F, message = F, warning = F)

library(tidyverse)
library(sf)
library(mapview)
library(plotly)
library(shiny)
library(leaflet)
library(kableExtra)

prj <- "+proj=utm +zone=11 +datum=NAD83 +units=m +no_defs"

source('R/funcs.R')

data(biodat)
data(rchdat)
```

This map shows the stream network and locations with biological data.  Select a species to view locations and tabular information of where and when the species was observed.

Add selector for source

```{r mapinp}
# select species to show
selectInput("spmap", "Choose species:", sort(unique(biodat$name)))
```

```{r reactives}
# species data to plot, table
sprct <- reactive({
  
  # input
  spmap <- input$spmap
  
  out <- biodat %>% 
    filter(name %in% spmap)
  
  return(out)
  
})

# map of species data
spmap <- reactive({
  
  # input
  sprct <- sprct()

  # get colors, pt size
  occol <- sprct$occol
  occex <- sprct$occex
  
  # remove irrelevant columns
  tomap <- sprct %>% 
    select(-occol, -occex, -FID_1)

  out <- mapview(rchdat, homebutton = F, color = 'grey') + 
    mapview(tomap, zcol = 'occurrence', col.region = occol, cex = occex, legend = T, homebutton = F, layer = 'occurrence')
 
  return(out)
 
})

# get number of sites sampled by year for a species
yrsmp <- reactive({
  
  # input
  sprct <- sprct()
  st_geometry(sprct) <- NULL

  # full year range
  yrs <- range(sprct$yr, na.rm = T)
  yrs <- seq(yrs[1], yrs[2])
  yrs <- data.frame(yr = yrs)
  
  # summarize  
  out <- sprct %>%
    select(name, yr, FID_1) %>% 
    na.omit %>% 
    unique %>% 
    group_by(yr) %>% 
    mutate(count = n()) %>% 
    ungroup %>% 
    select(yr, count) %>% 
    unique %>% 
    left_join(yrs, ., by = 'yr') %>% 
    tidyr::complete(yr)
  
  return(out)
  
})
# get number of sites where species was present/absent complete cases
yrcnt <- reactive({
  
  # input
  sprct <- sprct()
  st_geometry(sprct) <- NULL
  
  # full year range, by occurrence
  yrs <- range(sprct$yr, na.rm = T)
  yrs <- seq(yrs[1], yrs[2])
  yrs <- crossing(
    yr = yrs, 
    occurrence = c('present', 'absent')
  )

  # summarize
  out <- sprct %>%
    select(name, yr, occurrence) %>% 
    na.omit %>% 
    group_by(yr, occurrence) %>% 
    summarise(count = n()) %>% 
    ungroup %>% 
    left_join(yrs, ., by = c('yr', 'occurrence')) %>% 
    tidyr::complete(yr, occurrence)
  
  return(out)
  
})
```

```{r spmap}
output$spmap = renderLeaflet({spmap()@map})
leafletOutput('spmap', height = 500)
```

<br></br>
Counts of presence/absence by year for `r renderText(input$spmap)`:
```{r yrcnt, out.width = "100%"}
renderPlot({

  # inputs
  yrcnt <- yrcnt()
  
  out <- ggplot(yrcnt, aes(x = yr, group = occurrence, fill = occurrence, y = count)) + 
    geom_line() +
    geom_point(pch = 21, size = 5) +
    theme_bw(base_size = 16) +
    theme(
      axis.title.x = element_blank(), 
      axis.text.x = element_text(angle = 45, hjust = 1), 
      legend.position = 'top', 
      legend.title = element_blank()
    ) + 
    scale_fill_manual(values = c('tomato1', 'lightgreen')) +
    scale_y_continuous('Number of locations') + 
    scale_x_continuous(breaks = unique(yrcnt$yr))
  
  return(out)

})
```

Number of locations sampled by year for `r renderText(input$spmap)`:
```{r yrsmp, out.width = "100%"}
renderPlot({

  # inputs
  yrsmp <- yrsmp()

  out <- ggplot(yrsmp, aes(x = yr, y = count)) + 
    geom_line(colour = 'black') +
    geom_point(pch = 21, size = 5, fill = 'grey') +
    theme_bw(base_size = 16) +
    theme(
      axis.title.x = element_blank(), 
      axis.text.x = element_text(angle = 45, hjust = 1), 
      legend.position = 'top', 
      legend.title = element_blank()
    ) + 
    scale_y_continuous('Number of locations') + 
    scale_x_continuous(breaks = unique(yrsmp$yr))
  
  return(out)

})
```

Tabular summary

Downloads

